---
title: readnet
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

The goal of readnet is to provide tools for preprocessing network data (a companion to igraph, statnet, netdiffuseR, etc.)


Some of the features:

*   Automatically handle coding for nodes, time and groups.

*   If either `group.id` or `time` are specified, the function makes recursive calls, returning nested edgelists with the following structure:
    
    ```
    object
    --- group1:
      --- edgelist in time1
      --- ...
      --- edgelist in time n
    ...
    --- group n:
      --- edgelist in time1
      --- ...
      --- edgelist in time n
    ```
    
*   Includes functions to coerce the resulting edgelists into igraph objects and viceversa.

This project is on development and not published on CRAN yet.

## Installation

You can get `readnet` using devtools:

```r
devtools::install_github("USCCANA/readnet")
```

## Example

This is a basic example which shows you how to solve a common problem. Suppose that you have a dataset that looks something like this:

```{r fakedata, echo=FALSE}

dat0 <- data.frame(
    ego    = c("a", "b", "c"),
    alter1 = c("b", NA, NA),
    alter2 = NA,
    time   = 0
  )

dat1 <- data.frame(
  ego    = c("a", "b", "c"),
  alter1 = c("b", "a", NA),
  alter2 = c("c", NA, NA),
  time   = 1
  )

dat <- rbind(dat0, dat1)
```

```{r}
dat
```

This is, a survey with network nominations, and in this case, including a time variable. We can use the function `survey_to_edgelist` to process this data


```{r example}
library(readnet)
(ans <- survey_to_edgelist(dat, "ego", c("alter1", "alter2"), time = "time"))
```

The previous returned a list with elements of class `rn_edgelist`. Each element contains an edgelist, a vector of labels, and a dataframe with data. We can use the function `as_igraph` to turn this into an igraph object

```{r}
as_igraph(ans)
```

and as a `rn_edgelist` back

```{r}
as_rn_edgelist(as_igraph(ans))
```


